---
title: Red Wine Quality
author: Alex Trejo
date: "October 2015"
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
library(knitr)
library(gridExtra)
library(GGally)
library(psych)
library(car)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
if (!file.exists("wineQualityReds.csv")) {
  URL <- "https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityReds.csv"
  download.file(URL, destfile = "wineQualityReds.csv", method="curl")
}
df = read.csv("wineQualityReds.csv")
```

```{r echo=FALSE, Help_Functions}

# Reset par values to default ones
# Use: par(resetPar()) 
# Source: http://stackoverflow.com/a/5790430
resetPar <- function() {
    dev.new()
    op <- par(no.readonly = TRUE)
    dev.off()
    op
}
```

# <a id="Data"></a>Exploring the data
First of all, I want to Check the data dowloaded...
```{r Get_Dimensions}
dim(df)
```
check for missing values...
```{r Check_NA}
anyNA(df)
```
check for out of scale quality values...
```{r Check_quality_values}
any(df$quality > 10 || df$quality < 0)
```
check for negative values...
```{r Check_negative_values}
any(df < 0)
```

Once checked let's start with the analysis:

```{r echo=FALSE, Get_Vars_and_Types}
str(df)
```

There are 1599 different wines observations with 13 variables each one with the 
following distribution:

- 11 variables are chemical properties (independent variables)
- 1 variable, `X`, is the observation index (wine)
- 1 variable, `quality`, is the experts quality rate (dependent variable)

```{r echo=FALSE, Summary}
summary(df)
```

The general summary shows some interesting values to consider:

- The median `quality` is **6** in a 0 to 10 scale... 
- *Maximum* values for `total.sulfur.dioxide`, `volatile.acidity`, `citric.acid`,
`chlorides`, `residual.sugar` or `free.sulfur.dioxide` are between 4 and 7 times
the respective median

Given these extreme values, it will be interesting to see dispersion of the 
chemical properties:

```{r, echo=FALSE, Dispersion}
sapply(df[2:13], sd)
```

The `total.sulfur.dioxide`, `free.sulfur.dioxide` or `residual.sugar` dispesion 
is too high, as we suspected above.

# Univariate Plots Section

<a id="Histograms"></a>
How the variables are distributed?
Let's plot the histogram of each variable and see if they are normally distributed or not...

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=10, Histograms_Matrix}
# Create or empty list
pltList <- list()
# for each property in the df Datafreme starting in 2 (after the index X)
for (i in 2:length(df)) {
  
  # Binwith must be dynamic, so let's define best value for each property
  # Compute a sequence of about n+1 equally spaced ‘round’ values
  breaks <- pretty(range(df[,i]), n = nclass.scott(df[,i])/2, min.n = 1)
  
  # Compute the difference between the break values
  bwidth <- breaks[2]-breaks[1]
  
  # Store into the list of plots the ciorent plot
  pltList[[i-1]] <- ggplot(data = data.frame(value = df[,i]), 
                           aes(value, ..density..)) + 
    geom_histogram(aes(fill = ..count..), 
                   binwidth = bwidth) +
    geom_density() +
    scale_fill_gradient('Count', 
                        low = 'green', 
                        high = 'red') +
    xlab(colnames(df)[i]) +
    theme(legend.position='top')
}

# Once ended the loop arrange all the plots together
do.call(grid.arrange, c(pltList, list(nrow=3, ncol=4)))
```

Looking at the plots we can observe several properties right skewed, and few 
properties seems normally distributed like `pH`, `density` and `quality` as well, 
but the density line in their histogram is not showing that. let's plot the Normal Q-Q for each one of the variables 

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=10, QQ_Matrix}
# Setting Array Rows/Cols 
par(mfrow=c(3, 4))

# for each property in the df Datafreme starting in 2 (after the index X)
for (i in 2:length(df)) {
  
  # Plot the Normal Q-Q Plot 
  qqnorm(df[,i], main = paste('Normal Q-Q', colnames(df)[i]))
  
  # Plot the Normal Line in Blue color 
  qqline(df[,i],col = 'blue')
} 

```

The Normal Q-Q plot, confirms the same as we observed in the histograms.

Can we transform some of the variables an see if they fit a normal distribution?

<a id="Transformation"></a>
Yes, we can. Let's take a look what if we change the scale to logaritmic 
to all the vars...

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=10, Histograms_Matrix_Log}
# Create or empty list
pltList <- list()

# for each property in the df Datafreme starting in 2 (after the index X)
for (i in 2:length(df)) {
  
  # Binwith must be dynamic, so let's define best value for each property
  # Compute a sequence of about n+1 equally spaced ‘round’ values
  breaks <- pretty(range(df[,i]), n = nclass.scott(df[,i]), min.n = 1)
  
  # Compute the difference between the break values
  bwidth <- breaks[2]-breaks[1]
  
  # Store into the list of plots the current plot
  pltList[[i-1]] <- ggplot(data = data.frame(value = df[,i]), 
                           aes(value, ..density..)) + 
    geom_histogram(aes(fill = ..count..), binwidth = bwidth) +
    geom_density() +
    scale_x_log10() +
    scale_fill_gradient('Count', low = 'green', high = 'red') +
    xlab(colnames(df)[i]) +
    theme(legend.position = 'top')
}

# Once ended the loop arrange all the plots together
do.call(grid.arrange, c(pltList, list(nrow=3, ncol=4)))
```

Interesting... we can see that the variables skewed like `fixed.acidity`, 
`volatile.acidity`, `chlorides` and `sulphates` turn to a 'more' normal 
distribution, so Let's add them to the dataframe:

```{r Transforming_vars}

#adding transformed vars to the dataframe with the suffix '.log'
df$fixed.acidity.log <- log10(df$fixed.acidity)
df$volatile.acidity.log <- log10(df$volatile.acidity)
df$chlorides.log <- log10(df$chlorides)
df$sulphates.log <- log10(df$sulphates)
```

How Normal Q-Q plot loks like now?

Let's perform again a Normal Q-Q test with the new log variables:

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=10, QQ_Matrix_Log}
# Setting Array Rows/Cols 
par(mfrow=c(3, 4))

# Define a list with the Study Properties 
# (new log variables replacing the original ones)
studyProp <- c('fixed.acidity.log', 
               'volatile.acidity.log', 
               'citric.acid', 'residual.sugar', 
               'chlorides.log', 'free.sulfur.dioxide', 
               'total.sulfur.dioxide', 
               'density', 
               'pH', 
               'sulphates.log', 'alcohol', 'quality')

# Create a new Dataframe with only the Study Properties
dfStudy <- df[names(df) %in% studyProp]

# For each property in the Daytaframe
for (i in 1:length(dfStudy)) {
  
  # Plot the Normal Q-Q Plot 
  qqnorm(dfStudy[,i], main = paste('Normal Q-Q', colnames(dfStudy)[i]))
  
  # Plot the Normal Line in Blue color 
  qqline(dfStudy[,i], col = 'blue')
} 

```

Well... As seen before, it seems that the new variables perform better than 
the original ones. We will use them transformed and then we will get back to the original value.

Is there a variable more important than the others?

<a id="Normality"></a>
Yes, there is. Due to the importance of `quality` in this study, lets take a 
look at `quality` histogram deeper.

```{r echo=FALSE, Quality_Histogram}
ggplot(data = df, 
       aes(x = quality)) + 
  geom_histogram(aes(y = ..density.., fill = ..count..), 
                 binwidth = 1) +
  scale_fill_gradient('Count', low = 'green', high = 'red') +
  stat_function(fun = dnorm, 
                color = 'blue', 
                arg = list(mean = mean(df$quality), 
                           sd = sd(df$quality))) +
  geom_vline(xintercept  = median(df$quality), 
             size = 1.5, 
             linetype = 3) +
  scale_x_discrete(breaks = seq(0, 10, 1)) +
  coord_cartesian(c(0,10)) +
  xlab('Wine quality')
```

Some insights about `quality`:

- Most of wines are rated between 5 and 6, with a median of 6 while mean is 5.636
- No wines rated below 3 or upper 8
- Distribution seems normal

Lets test the normality of `quality`

```{r echo=FALSE, Normality_tests}
shapiro.test(df$quality)

# Setting Array Rows/Cols 
par(mfrow=c(1, 1))

# Plot the Normal Q-Q Plot
qqnorm(df$quality)

# Plot the Normal Line in Blue color
qqline(df$quality, col = "blue")
```

Well... looking at the resutls we can NOT ensure that `quality` is normally 
distributed, but we can assume it with a relative high level of confidence, 
considering the sample size and the normal curve. The Q-Q plot, support my 
decision, taking into account that `quality` is a discrete variable. 

# Univariate Analysis

### What is the structure of your dataset?
See [Exploring the Data](#Data)

### What is/are the main feature(s) of interest in your dataset?
The main feature of interest is `quality` as discused [here](#Normality)

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

At this point is difficult to answer this question, but intuitively I suggest 
`alcohol`, acidity related features,  and sulfur Dioxide, but not sure which 
one of the two `free.sulfur.dioxide` or `total.sulfur.dioxide`. I need a 
correlation test to answer this question with more confidence, that I'll 
performe in the next section.

### Did you create any new variables from existing variables in the dataset?
Yes, I did. I eplain why in the next question. Additionally, I'm thinking to 
split `quality` into `bad`, `medium` and `good` wines 

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

As we can see in the [Features Histograms & Normality](#Histograms),
`citric.acid` is the feature with an unusual distribution.`residual.sugar` 
and `chrolides` have a right-skewed distribution with a heavy long tail, 
what sugests the existence of outliers. 

I've performed a logaritmic [transformation](#Transformation) to 
`fixed.acidity`, `volatile.acidity`, `chlorides` and `sulphates` due to their 
right-skewed distribution, that transforms the var into a more normalized 
distribution.

# Bivariate Plots Section

Let's start with the correlation matrix applied to the dataframe with the new 
variables:

```{r echo=FALSE, Bivariate_Plots}
# Correlation Test with Plots... ¡¡5-10 min of compute time!!
# ggpairs(dfStudy)

# Correlation Test Array values
cor(dfStudy)
```

Analyzing the correlations, we can observe that the most correlated properties 
with `quality` are `fixed.acidity.log`, `volatile.acidity.log`, `citric.acid`, 
`chlorides.log`, `total.sulfur.dioxide`, `density`, `sulphates.log` 
and `alcohol`.

How is the relation of quality with the rest of the main properties?

```{r echo=FALSE, DataFrame_with_correlated_vars}

# Define a list with the Main Correlated Properties
corProp <- c('fixed.acidity.log', 
             'volatile.acidity.log', 
             'citric.acid', 
             'chlorides.log', 
             'total.sulfur.dioxide', 
             'density', 
             'sulphates.log', 
             'alcohol', 
             'quality')

# Create a new Dataframe with only the Main Correlated Properties
dfCor <- df[names(df) %in% corProp]

# Reorder Dataframe with quality [5] in the last index
dfCor <- dfCor[c(1,2,3,4,6,7,8,9,5)]
```


```{r echo=FALSE, fig.width=10, Quality-Property_boxplots}
# Create or empty list
pltList <- list()

# for each property in the df Datafreme do
for (i in 1:8) {
  
  # Store into the list of plots the current plot
  pltList[[i]] <- ggplot(aes_string(x = 'quality', 
                                    y = colnames(dfCor)[i]), 
                         data = dfCor) + 
    geom_boxplot(aes_string(group = 'quality')) +
    geom_line(color = 'blue', 
              stat = 'summary', 
              fun.y = median) +
    ylim(quantile(dfCor[,colnames(dfCor)[i]], 0.05), quantile(dfCor[,colnames(dfCor)[i]], 0.95)) +
    scale_x_continuous(limits = c(2,9), 
                       breaks = seq(3,8,1))
}

# Once ended the loop arrange all the plots together
do.call(grid.arrange, c(pltList, list(nrow=2, ncol=4)))
```

Well, using boxplots, grouped by quality and ploting the line that joins 
the medians of each group, we can see the relation closely, by looking at the 
slope of this line. Some comments about these plots:

- The `density`, `total.sulfur.dioxide`, `chlorides.log`and `sulphates.log` 
properties have an several outliers in their plots (dots outside the box plot 
tails)
- The `citric.acid`, `fixed.acidity.log` and `alcohol`, are the properties with
largest interquartilic range (IQR) while `chlorides.log` is the one with the 
smallest IQR

Is it possible to group quality into a more simple variable? Yes, it is... 
I will create a new factorized variable called `quality.bucket`
as I mentioned abobe, by cutting the original `quality` variable into `bad`, 
`medium` and `good` wines,. The cutting levels will be:

- `bad` if `quality` is <= 4
- `medium` if `quality` > 4 and <= 6
- `good` if `quality` is > 6

```{r Cutting_quality}
# Create a new Factorized variable cuting the quality original variable
dfCor$quality.bucket <- cut(dfCor$quality, 
                            breaks = c(0, 4, 6, 10),
                            right = TRUE, 
                            labels = c('Bad', 'Medium', 'Good'))
```

Using the same plotting technique as above, Let's see the boxplots now:

```{r echo=FALSE, fig.width=10, Quality.bucket-Property_boxplots}
# Create or empty list
pltList <- list()

# for each property in the df Datafreme do
for (i in 1:8) {
  
  # Store into the list of plots the current plot
  pltList[[i]] <- ggplot(aes_string(x = 'quality.bucket', 
                                    y = colnames(dfCor)[i], 
                                    group=1), 
                         data = dfCor) + 
    geom_boxplot(aes_string(group = 'quality.bucket')) +
    geom_line(color = 'blue', 
              stat = 'summary', 
              fun.y = median) +
    ylim(quantile(dfCor[,colnames(dfCor)[i]], 0.05), quantile(dfCor[,colnames(dfCor)[i]], 0.95)) 
}

# Once ended the loop arrange all the plots together
do.call(grid.arrange, c(pltList, list(nrow=2, ncol=4)))
```

Well.. Now is easy to see that tne line are smoother, due tho the reduction 
of levels in quality (3 now 6 before). Some comments:

- The outliers are concentrated basically in `Medium` group, due essentially 
because is the group with more observations inside. 
- The slope in the medians line is clarly showing the relation excepting in 
`total.sulfur.dioxide`

How many observations contains each group?

```{r, echo=FALSE, Observations_Table}
dfCor.by_quality.bucket <- dfCor %>%
  group_by(quality.bucket) %>%
  summarise(n = n())
kable(dfCor.by_quality.bucket)
```

As I commented before, the group `Medium` concentrates the most observations.

Which are the median for each group and each property?

```{r, echo=FALSE, Meadians_Table}
dfCor.median_by_quality.bucket <- dfCor %>%
  group_by(quality.bucket) %>%
  summarise_each(funs(median))
kable(dfCor.median_by_quality.bucket)
```

The `total.sulfur.dioxide` median values for each group show what we have seen
with the slope of the median lines, that the median values for each group 
vary wwith no clear pattern... from 26 to 40 and then down to 27 again...

On the other hand, medians for other properties like `citric.acid` are grater
leval by level. Special mention to `alcohol` where the median for `good` wines
is 11.6 % of volume while `bad` or `Medium` wines is the same, 10. 

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

In this part of the study, I selected few variables (`fixed.acidity.log`, 
`volatile.acidity.log`, `citric.acid`, `chlorides.log`, `total.sulfur.dioxide`, 
`density`, `sulphates.log` and `alcohol`), that seems to have more correlation 
with ths `quality` variable, thanks to the correlation matrix.

After taht, I decided to plot each of the selected variables using the boxplot, 
getting some interesting insights:

- `citric.acid`, `alcohol`, `fixed.acidity.log` and `sulphates.log` have 
**positive** relation with the `quality` of wine.
- `density`, `volatile.acidity.log` and `chlorides.log` have **negative** 
relation with the `quality` of wine
- `total.sulfur.dioxide` has **no aperent** relation

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

There are several interesting realtions that need to be observed:

- As I can expect, `pH` has strong correlation with the acidity rtelated 
properties like `acid.citric` (positive)  or `fixed.acidity` (positive) .
- `density` is correlated with `acid.citric`  or `residual.sugar`, as well as 
`alcohol` or `fixed.acidity`. The realtion of the `density` with the 
`residual.sugar` (positive) or `alcohol` (negative) is expected, but not with 
the `acid.citric` (positive) or the `fixed.acidity` (positive).

### What was the strongest relationship you found?

`quality` is strongly correlated with `alcohol` near followed by `volatile.acidity`. 

# Multivariate Plots Section

Let's see how this main features are related between them (all against all), 
but now grouped by `quality.bucket`:

```{r echo=FALSE, fig.width=10, fig.height=10, ScatterplotMatrix}
#par(bty="l",las=2)
#par(resetPar())
par(xpd=FALSE)
scatterplotMatrix(~citric.acid + 
                    fixed.acidity.log + 
                    volatile.acidity.log + 
                    density + 
                    total.sulfur.dioxide + 
                    sulphates.log + 
                    alcohol | quality.bucket, 
                  data = dfCor, 
                  legend.plot = F)
par(xpd=TRUE)
legend(0.9, 1.06,
       legend = levels(dfCor$quality.bucket),
       inset = c(0.2,0),
       pch = c(1,2,3), 
       col = c(1,2,3),
       bty = 'n',
       cex = 0.7)
```

Here we can see most of the things seen before, plus some new interesting relations:

- Strong positive correlation between `density` and `fixed.acidity.log`,
`fixed.acidity.log` and `citric.acid`, or `sulphates.log` and `citric.acid`
- Strong negative correlation between `density` and `alcohol` or 
`volatile.acidity.log` and `citric.acid`

Let's take a look more closely at how `alcohol` is realted with the other 
main variables grouped by `quality.bucket` and plot the density curves, and 
the average for each group:

```{r echo=FALSE, Reordering_dfCor}
# Reorder Dataframe with Alcohol [4] before quality [9]
dfCor <- dfCor[c(1,2,3,5,6,7,8,4,9,10)]
```

```{r, echo=FALSE, Means_Table}
dfCor.mean_by_quality.bucket <- dfCor %>%
  group_by(quality.bucket) %>%
  summarise_each(funs(mean))
```

```{r echo=FALSE, fig.width=10, fig.height=30, Multivariate_Plots}
# Create or empty list
pltList <- list()

# for each property in the df Datafreme do
for (i in 1:7) {
  
  # Store into the list of plots the current plot
  pltList[[i]] <- ggplot(aes_string(x = 'alcohol', 
                                    y = colnames(dfCor)[i], 
                                    colour = 'quality.bucket'), 
                         data = dfCor) +
    stat_density2d(aes(alpha = ..level..)) +
    ggtitle(paste(colnames(dfCor)[i], 'vs Alcohol')) +
    geom_point(data = dfCor.mean_by_quality.bucket, 
              aes_string(x = 'alcohol',
                         y = colnames(dfCor)[i], 
                         colour = 'quality.bucket',
                         size = 5)) +
    scale_size(guide = 'none') +
    geom_line(data = dfCor.mean_by_quality.bucket, 
              aes_string(x = 'alcohol',
                         y = colnames(dfCor)[i],
                         group = 1),
              color = 'black',
              linetype = 3) +
    theme(legend.position='left')
}

# Once ended the loop arrange all the plots together
do.call(grid.arrange, c(pltList, list(nrow=7, ncol=1)))
```

These series of plots, give us a map of how the observations are distributed
for each `quality.bucket` and the density level for each one. In general, we can 
observe that most of the means have some correlation, increasing in each pair 
by `quality.bucket`... In the next section I explain some findings.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

Using Multivariate Analysis, let me observe more closely the relations between 
the properties, and some of the relations found at Bivariate Analysis 
has been confirmed:

- `volatile.acidity` and `density` has strong negative correlation with 
`quality` and `alcohol` 
- `alcohol` has strong positive correlation with `quality`, `citric.acid` 
and `sulphates`

### Were there any interesting or surprising interactions between features?

Yes, there were. First of all, that the amount of `alcohol` in a wine, seems 
to increase the `quality`, and second, that the *good* wines tend to have 
more `sulphates`, which change my idea that 'the more preservatives a wine 
has worst is the quality'. 

------

# Final Plots and Summary

For this part of the Study, I will transform the log variables to the original 
ones, so the values shown in the plots will be the correct.

```{r echo=FALSE, Final_Properties_DataFrame}
# Selecting the Most Important properties
finalProp <- c('volatile.acidity.log', 
               'citric.acid', 
               'sulphates.log', 
               'alcohol', 
               'quality', 
               'quality.bucket')

# Create new Dataframe with only the Important Properties
dfFinal <- dfCor[names(dfCor) %in% finalProp]

# Get back to original values the properties transformed
dfFinal$volatile.acidity.log <- exp(dfFinal$volatile.acidity.log)
dfFinal$sulphates.log <- exp(dfFinal$sulphates.log)

# Renaming columns
colnames(dfFinal) <- c('Citric_Acid', 
                       'Volatile_Acidity', 
                       'Sulphates', 
                       'Alcohol', 
                       'Quality_Rate', 
                       'Qualification')

# Define units for each property
units <- list("(g/dm^3)", "(g/dm^3)", "(g/dm^3)", "(% Volume)")
```

### Plot One
```{r echo=FALSE, fig.width=10, Plot_One}
# Create or empty list
pltList <- list()

# for each property in the df Datafreme do
for (i in 1:4) {
  
  # Store into the list of plots the current plot
  pltList[[i]] <- ggplot(aes_string(x = 'Qualification', 
                                    y = colnames(dfFinal)[i], 
                                    group=1), 
                         data = dfFinal) +
    geom_point(aes_string(color = 'Qualification', alpha = 1/20), 
               position = position_jitter(),
               size = 2) +
    geom_boxplot(aes_string(group = 'Qualification', alpha = 0)) +
    stat_smooth(size = 1, 
                colour = 'red', 
                method = 'lm',
                fullrange = TRUE) +
    scale_alpha(guide = 'none') +
    ylim(quantile(dfFinal[,colnames(dfFinal)[i]], 0.05), quantile(dfFinal[,colnames(dfFinal)[i]], 0.95)) +
    ggtitle(paste(gsub("_", " ",colnames(dfFinal)[i]), 'vs Wine quality')) +
    ylab(paste(gsub("_", " ",colnames(dfFinal)[i]), units[[i]])) +
    guides(color = F)
}

# Once ended the loop arrange all the plots together
do.call(grid.arrange, c(pltList, list(nrow=2, ncol=2)))
```

### Description One

This plot is interesting because concentrates the main features that, 
after my investigation, seems to influence more in the wine `quality`. 

- Three of them are positive correlated `citric.acid`, `sulphates.log` and `alcohol`. 
- The remaining one is `volatile.acidity` which is negative correlated with `quality`.

### Plot Two
```{r, echo=FALSE, Means_Final_Table}
dfFinal.mean_by_qualification <- dfFinal %>%
  group_by(Qualification) %>%
  summarise_each(funs(mean))
```
```{r echo=FALSE, fig.width=10, Plot_Two}
ggplot(aes(x = Alcohol, 
           y = Volatile_Acidity, 
           colour = Qualification), 
       data = dfFinal) +
  stat_density2d(aes(alpha = ..level..)) +
  geom_point(data = dfFinal.mean_by_qualification, 
              aes(x = Alcohol,
                  y = Volatile_Acidity, 
                  colour = Qualification,
                  size = 5)) +
  geom_line(data = dfFinal.mean_by_qualification, 
            aes(x = Alcohol,
                y = Volatile_Acidity,
                group = 1),
                color = 'black',
                linetype = 3) +
  geom_smooth(data = dfFinal.mean_by_qualification,
              aes(x = Alcohol,
                  y = Volatile_Acidity,
                  group = 1),
              method=lm, 
              fullrange=TRUE,
              fill = NA,
              color = 'purple') +
  scale_size(guide = 'none') +
  theme(legend.position = 'left') +
  ggtitle('Alcohol vs Volatile Acidity by wine quality') +
  xlab('Alcohol (% Volume)') +
  ylab('Volatilile Adidity (g/dm^3)') +
  theme_classic() +
  coord_cartesian(ylim = c(0.5, 1.1), xlim = c(8, 14))
```

### Description Two

This is one of the most important conclusions extracted from my investigation. 
The Quality of wine gets better with higher levels of alcohol and lower 
Volatile Acidity. In Purple, we can see the smoothed line for the three means
corresponding to each group of qualifications showing a clear negative 
correlation. Despite the similarity of % of Alcohol in the `Bad`and `Medium` 
groups, we can observe the change in the `Good` group.

### Plot Three
```{r echo=FALSE, fig.width=10, Plot_Three}
ggplot(aes(x = Quality_Rate, 
           y = Sulphates,
           fill = Qualification, 
           group=1), 
       data = dfFinal) + 
  geom_boxplot(aes(group = Quality_Rate)) +
  stat_smooth(size = 1, 
              colour = "red", 
              method = 'gam') +
  ggtitle('Sulphates vs Quality') +
  labs(y = 'Sulphates (g/dm^3)',
       x = 'Quality') 
```

### Description Three

The last plot shows the second important conclusion. The Quality of wine gets 
better with higher levels of sulphates. I supose that with sulphates, as a 
good preservative, the wine gets worst slowly than the more 'natural' ones. 
More Sulphates and more Alcohol, better Wine.

------

# Reflection

The Exploratory Data Analysis done in this project and along all the course, 
has been very useful to understand not only the utilization of R, but for 
understanding meaning of some staistical tools in real situations. In this 
case, the quality of red wines in function of other chemical properties. 
Tools like scatterplots or density plots, helped me to draw the long list of 
values in a simple and meaningful way.

I had some struggles with R... like how to loop over each property and plot them 
together in a simple way. Inside these loops, how to label each property
with its corresponding units, etc.. I tried to avoid repetitive code, and the use
of loops helped me a lot, but this solution has inherit problems which I had to 
deal with... after fighting a lot, the plots were appearing in the way I imagined,
so that's for me my biggest success.

The findings were surprisingly positive, like the positive relation between 
alcohol and quality, or the negative one with amount of sulphates. Before this 
study, I was completely misunderstood, because I thought that a 'bad' wine had
more % of Alcohol. 

We need to keep in mind, that this is a *small* sample of red wines, and the 
quality variable is a subjective value of an expert, with all the implications 
that have... maybe a better solution is to have a median value of a group of 
experts.

The groups made to classify the quality of wines, are not balanced in terms 
of number of observations, so I consider that we will need more `bad` and `good` 
wines to strengthen this study.

An other consideration, is that it would have been interesting to have the 
Geolocation of each observation, so we could compare the wine quality with 
the designation of origin, and plot it into a map. The result could be 
interesting...

# References

- http://stackoverflow.com/questions/5319754/cross-reference-named-anchor-in-markdown
- http://stackoverflow.com/questions/26235825/for-loop-only-adds-the-final-ggplot-layer
- http://stackoverflow.com/questions/14200027/how-to-adjust-binwidth-in-ggplot2
- http://stackoverflow.com/questions/29182228/plotting-normal-curve-over-histogram-using-ggplot2-code-produces-straight-line
- http://stackoverflow.com/questions/7781798/seeing-if-data-is-normally-distributed-in-r
- http://docs.ggplot2.org/current/aes_group_order.html
- http://stackoverflow.com/questions/15458526/r-pass-variable-column-indices-to-ggplot2
- http://www.monashscientific.com.au/VolatileAcidSignificance.htm
- http://stackoverflow.com/questions/5789982/reset-par-to-the-default-values-at-startup
- http://stackoverflow.com/a/24455439